---
title: "BMI_final_project"
author: "Yuqi Tu, Yue (Lynette) Pan, Nadiya Pavlishyn, Shumin Rui"
date: "December 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(dplyr)
library(forcats)
library(leaps)
```

```{r}
#####load in data
data = read_excel("GHProject_Dataset.xlsx") %>%
  janitor::clean_names() %>% 
  tidyr::separate(admitdtm, into = c("weekday", "admitdate"), sep = "day, ") %>%
  mutate(admitdate = gsub(", ", "/", admitdate)) %>% 
  tidyr::separate(admitdate, into = c("month", "dateyear"), sep = " ") %>% 
  mutate(month1 = match(month,month.name)) %>% 
  mutate(date = paste(stringr::str_sub(dateyear,-4,-1),"/",month1,"/",stringr::str_sub(dateyear,1,2),sep = ""),
         log_day = log(losdays2)) %>% 
  #filter out extreme temperatures
  filter(temperature < 48, temperature > 29) %>% 
  select(-month, -dateyear, -weekday, -month1) %>% 
  mutate(icu_flag = as.factor(icu_flag),
         is30dayreadmit = as.factor(is30dayreadmit))

#arrange the data according to the time of admission
data = data[order(data$date),] 

#remove those more than one visits
data = data[!duplicated(data$patientid),]
data = data[!(data$is30dayreadmit == 1 & data$losdays2 > 30),]
data = data %>%
  select(patientid, visitid, loshours, losdays2, date, cindex, evisit, icu_flag, ageyear, gender, race, religion, maritalstatus, facilityname, insurancetype, bpsystolic, o2sat, temperature, heartrate, respirationrate, bpdiastolic, log_day, is30dayreadmit) %>% 
#remove the observation with NAs
  na.omit()

#tidying: collapsing, setting indicator variables, 
#####descriptive statistics of each variable
##continuous variable
sum1 <- summary(data$losdays2,digits = 3) #transformation get good results 
sum1["Std.Dev."] <- round(sd(data$losdays2),3)

sum2 <- summary(data$ageyear,digits = 3) 
sum2["Std.Dev."] <- round(sd(data$ageyear),3)

sum3 <- summary(data$bpsystolic,digits = 3) 
sum3["Std.Dev."] <- round(sd(data$bpsystolic),3)

sum4 <- summary(data$o2sat,digits = 3)
sum4["Std.Dev."] <- round(sd(data$o2sat),3)

sum5 <- summary(data$temperature, digits = 3)
sum5["Std.Dev."] <- round(sd(data$temperature),3)

sum6 <- summary(data$heartrate, digits = 3)
sum6["Std.Dev."] <- round(sd(data$heartrate),3)

sum7 <- summary(data$respirationrate, digits = 3)
sum7["Std.Dev."] <- round(sd(data$respirationrate),3)

sum8 <- summary(data$bpdiastolic, digits = 3)
sum8["Std.Dev."] <- round(sd(data$bpdiastolic),3)

knitr::kable(dplyr::bind_rows(
  c(variable = "Losdays2",sum1),#good to do log transform
  c(variable = "AgeYear", sum2),#quite normal
  c(variable = "bpsystolic",sum3),
  c(variable = "o2sat", sum4), # 10 observation below 90, 80 observation above 100, 2% of the data
  c(variable = "temperature", sum5), #3 less than 30 and 2 more than 50
  c(variable = "heartrate", sum6), # 2 more than 240 (rare heart rate or caused by disease)
  c(variable = "respirationrate", sum7),
  c(variable = "bpdiastolic", sum8),
))

##categorical variable
#MEWS: The Modified Early Warning Score (MEWS) 
#0-1=normal, 2-3=increase caution, 4-5=further deterioration, >5 immediate action required
#table(data$mews)
#make them into 3 category and try to make the data balanced (though not perfect)
#data$mews_catg<-ifelse(data$mews==0|data$mews==1,"normal",
                       #ifelse(data$mews==2|data$mews==3,"increase caution",
                       #ifelse(is.na(data$mews) == TRUE,NA,"fur deter or worse")))
#data = data %>% mutate(mews_catg = fct_relevel(mews_catg,"increase caution"))
#table(data$mews_catg)
#mews_catg: 1=normal; 2=increase caution; 3 = further deterioration or immediate action required

#Cindex -- Charlson comorbidity index (CCI) ranks patients based on severity of comorbidity: 0=normal, 1-2=mild, 3-4=moderate and >5=severe
table(data$cindex)
#make them into three category to balance the data
data$cindex_catg<-ifelse(data$cindex==0,"normal",
                         ifelse(data$cindex==1|data$cindex==2,"mild",
                         ifelse(is.na(data$cindex)==TRUE,NA,"moderate to severe")))
data = data %>% mutate(cindex_catg = fct_relevel(cindex_catg,"mild"))
table(data$cindex_catg)

#evisit
table(data$evisit)
data$evisit_catg<-ifelse(data$evisit==0|data$evisit==1,"No or1 visit",
                  ifelse(is.na(data$evisit)==TRUE,NA,"More than 1"))
data = data %>% mutate(evisit_catg = fct_relevel(evisit_catg,"No or1 visit"))
table(data$evisit_catg)
#evisit_catg:1=0 or 1 emergency department visit;2=more than once emergency department visit(may indicate the serious sickness)

#ICU_Flag: 1=if during hospitalization, the patient had a visit in the intensive care unit (ICU); 0=otherwise. 
table(data$icu_flag)
#gender: patient???s gender -- balanced
table(data$gender)
#race: patient???s race -- quite unbalanced
table(data$race)
data$race_catg<-ifelse(data$race=="White","White",
                ifelse(data$race=="African Amer/Black","African Amer/Black","Other race"))
data = data %>% mutate(race_catg = fct_relevel(race_catg, "White"))
table(data$race_catg)
#put other race together, might be controversial, but is more balanced

#religion: patient's religion -- unbalanced, lets recode some
data$religion[data$religion == "Angelican"] <- "Christian"
data$religion[data$religion == "Non Denominational"] <- "Christian"
data$religion[data$religion == "Catholic"] <- "Christian"
data$religion[data$religion == "Hebrew"] <- "Jewish"
data$religion[data$religion == "Mormon"] <- "Other"
data = data %>% mutate(religion = fct_relevel(religion, "Christian"))
table(data$religion)


#MaritalStatus: patient's marital status-- unbalanced
table(data$maritalstatus)
data$marital_catg <- ifelse(data$maritalstatus=="Civil Union"|data$maritalstatus=="Married"|data$maritalstatus=="Separated","Not alone",
                      ifelse(data$maritalstatus=="Divorced"|data$maritalstatus=="Single"|
                             data$maritalstatus=="Widowed","Alone",NA))
table(data$marital_catg)

#InsuranceType: patient's insurance
table(data$insurancetype)
data$insurance_catg <- ifelse(data$insurancetype=="Private","Private",
                        ifelse(data$insurancetype=="Medicaid"|data$insurancetype=="Medicare","Public",NA))
table(data$insurance_catg)

#facility name
table(data$facilityname)



#visualize dataset for each variables
attach(data)
par(mfrow=c(3,3))
hist(losdays2) #highly skewed to the right
hist(log_day)
hist(ageyear)
hist(bpsystolic)
hist(o2sat)
hist(temperature)
hist(heartrate)
hist(respirationrate)
hist(bpdiastolic)

#try transformations of data
par(mfrow=c(1,1))
hist(data$losdays2) 
```
```

```{r}
## put (absolute) correlations on the upper panels,
## with size proportional to the correlations.
panel.cor <- function(x, y, digits = 2, prefix = "", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- abs(cor(x, y))
    txt <- format(c(r, 0.123456789), digits = digits)[1]
    txt <- paste0(prefix, txt)
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)*3
    text(0.5, 0.5, txt, cex = cex.cor * r)
}


#correlation matrix
pairs(~log(losdays2) + ageyear + bpsystolic + o2sat+ temperature +log(heartrate) +log(respirationrate)+ bpdiastolic, data=data, lower.panel = panel.smooth, upper.panel = panel.cor,main = "Correlation matrix for continuous data" )



### try some simple linear regression
#continuous variable
summary(lm(log(losdays2)~ageyear,data = data))
summary(lm(log(losdays2)~bpsystolic,data = data))
summary(lm(log(losdays2)~o2sat,data = data))
summary(lm(log(losdays2)~temperature,data = data))
summary(lm(log(losdays2)~log(heartrate),data = data))
summary(lm(log(losdays2)~log(respirationrate),data = data))
summary(lm(log(losdays2)~bpdiastolic,data = data))


#categorical varialbe
summary(lm(log(losdays2)~as.factor(cindex_catg),data=data))
summary(lm(log(losdays2)~as.factor(evisit_catg),data=data))
summary(lm(log(losdays2)~as.factor(icu_flag),data=data))
summary(lm(log(losdays2)~as.factor(gender),data=data))
summary(lm(log(losdays2)~as.factor(race_catg),data=data))
summary(lm(log(losdays2)~religion,data = data))
summary(lm(log(losdays2)~insurance_catg, data = data))
summary(lm(log(losdays2)~marital_catg, data = data))


#fit bmi again with 2 levels for BMI obese and not obese, nas not omitted
#data$bmi_catg_2<-ifelse(data$bmi<25,"not obese", "obese")
#summary(lm(log(losdays2) ~ bmi_catg_2, data = data))

#fit bmi with 2 levels for BMI obese and not obese, nas omitted
#data_bmiomit$bmi_catg_2<-ifelse(data_bmiomit$bmi<25,"not obese", "obese")
#summary(lm(log(losdays2) ~ bmi_catg_2, data = data_bmiomit))
```


```{r}
###build the main model
whole_model = lm(log_day ~ cindex_catg + evisit_catg + icu_flag + ageyear + gender + race_catg + religion + marital_catg +  insurance_catg + bpsystolic + o2sat + temperature + heartrate + respirationrate + bpdiastolic + is30dayreadmit + facilityname, data = data)
car::vif(whole_model)
###try automatic procedure
#backward procedure/stepwise regression based on AIC
step(whole_model, direction='backward')
step(whole_model, direction='both')
model_auto = lm(formula = log_day ~ cindex_catg + evisit_catg + ageyear + 
    gender + race_catg + marital_catg + insurance_catg + bpsystolic + 
    o2sat + temperature + heartrate + respirationrate + bpdiastolic + 
    is30dayreadmit + facilityname, data = data)
summary(model_auto)

#stepwise regression/backward procedure based on p-value from SAS (threshold p-value = 0.15)
model_sas = lm(log_day ~ ageyear + evisit_catg + bpsystolic + heartrate + cindex_catg + is30dayreadmit + insurance_catg + respirationrate + temperature + bpdiastolic + gender + marital_catg + race_catg + facilityname + o2sat, data = data)
summary(model_sas)

###criterion procedure
best <- function(model, ...) 
{
  subsets <- regsubsets(formula(model), model.frame(model), ...)
  subsets <- with(summary(subsets),
cbind(p = as.numeric(rownames(which)), which, rss, rsq, adjr2, cp, bic))
  return(subsets)
} 
# Select the 'best' 1 models of all subsets
round(best(model_sas, nbest = 1), 3)


###model selection
#We get the first model using automatic procedure
model1 = lm(log_day ~ ageyear + evisit_catg + bpsystolic + heartrate + cindex_catg + is30dayreadmit + insurance_catg + respirationrate + temperature + bpdiastolic + gender + marital_catg + race_catg + facilityname + o2sat, data = data)

summary(model1)
#noticing the model chose from criterion procedure, we get model 2
model2 = lm(log_day ~ ageyear + evisit_catg + bpsystolic + heartrate + cindex_catg + is30dayreadmit + insurance_catg + respirationrate , data = data)
summary(model2)

###compare the two models
anova(model2, model1)
# so model 1 is better
### drop the non-significant ones: race, facility name and o2sat
model1 = lm(log_day ~ ageyear + evisit_catg + bpsystolic + heartrate + cindex_catg + is30dayreadmit + insurance_catg + respirationrate + temperature + bpdiastolic + gender + marital_catg, data = data)
summary(model1)
final_model = model1
###model diagnostics
par(mfrow=c(2,2))
plot(final_model)

#possible leverages - no positive leaverage
#possible outliers in Y
stu_res = rstandard(final_model)
outliers_y1 = stu_res[abs(stu_res)>2.5]
data_drop1 = data[-c(8,43,139,144,199,224,247,294,346,748,763,777,1103,1112,1130,1138,1156,1157,1294,1439,1462,1523,1563,1585,1588,1634,1651,1881,2025,2073,2275,2320,2339,2456,2462,2520,2672,2685,2743,2775,2803,2820,2854,2935,2935,2946,2990,2993,3018,3068,3114,3148,3184,3456),]
model_drop1 =  lm(log_day ~ ageyear + evisit_catg + bpsystolic + heartrate + cindex_catg + is30dayreadmit + insurance_catg + respirationrate + temperature + bpdiastolic + gender + marital_catg, data = data_drop1)
summary(model_drop1)
```

```{r}
###cross validation/bootstrap
final_dataset = data %>%
  select(patientid,log_day,ageyear,evisit_catg,bpsystolic, heartrate, cindex_catg, is30dayreadmit, insurance_catg,  respirationrate, temperature, bpdiastolic, gender, marital_catg,race_catg, facilityname)
boot_sample = function(df){
  sample_frac(df,replace = TRUE)
}
boot_sample(final_dataset)
boot_straps = data_frame(
 strap_number = 1:2000,
 strap_sample = purrr::rerun(2000, boot_sample(final_dataset))
)
  
bootstrap_results = 
  boot_straps %>%
  mutate(models = purrr::map(strap_sample, ~lm(log_day ~ ageyear + evisit_catg + bpsystolic + heartrate + cindex_catg + is30dayreadmit + insurance_catg + respirationrate + temperature + bpdiastolic + gender + marital_catg, data = data)
  ),
         results = purrr::map(models, broom::tidy)) %>%
  select(-strap_sample, -models) %>%
  tidyr::unnest() %>%
  group_by(term) %>%
  summarize(mean_estimate = mean(estimate))
```

```{r}
## cook distance
cooksd = cooks.distance(model1)
plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")
abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")  # add labels
```


```{r}
## If we excluded the abnormal obervations
## temperature > 41 celsius

##filter_temp_bmi= data %>% 
##filter(temperature <=41 | bmi <= 55)


##full_model_with_interaction =lm(log(losdays2) ~ is30dayreadmit + cindex + evisit + ageyear + race_catg + insurance_catg + marital_catg + 
## o2sat + temperature + heartrate + respirationrate + bpdiastolic +  is30dayreadmit:cindex + is30dayreadmit:facilityname + is30dayreadmit:o2sat + evisit:ageyear + ageyear:marital_catg + gender:heartrate + gender:bpdiastolic + facilityname:insurance_catg + facilityname:respirationrate, data = filter_temp_bmi)


##step(full_model_with_interaction, direction='both')
```

